---
title: 'Homework #9'
author: "Colleen Yancey"
date: "March 21, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Structured Programming Practice
##Psuedo code
```{r}
# Call libraries needed
library(ggplot2)
library(reshape2)

# Create Variables
nGroup <- 4
nName <- c("Control", "Light", "Moderate", "Extreme") 
nSize <- c(10,10,10,10) # both are n = 6
nMean <- c(0.6, 0.75,0.95,1.1) # mean response in each group
nSD <- c(0.3,0.3,0.3,0.3) # variance in each group
ID <- 1:(sum(nSize))


```


```{r}
####################################
# Function: RandomData
# These function will create random normal data to help simulate and test hypotheses.
#input: mean, n, standard deviation
#output: simulated random normal data set
#-----------------------------------------
RandomData <-function(mean1=.6, mean2=.75,mean3=.95,mean4=1.1,n1=10,n2=10,n3=10, n4=10, sd1=.3, sd2=.3,sd3=.3,sd4=.3){
  myDF<-data.frame(Control=rnorm(mean=mean1, n=n1, sd=sd1), Light=rnorm(mean=mean2, n=n2, sd=sd2), Moderate=rnorm(mean=mean3, n=n3, sd=sd3), Extreme=rnorm(mean=mean4, n=n4, sd=sd4))
  myDF2<-melt(myDF)
  return(myDF2)
}
df<-RandomData()
################################
#Function: myANOVA
#The function will perform an ANOVA analysis and spit generate p values
#input: data frame from 1st function
#output: p values analysis
#-----------------------------
myANOVA <- function(dataFrame=df){
 ANOVA <- aov(value~variable, data=dataFrame)
  myOutput <- summary(ANOVA)
  return(myOutput)
}
##############################
#Function: ANOVAPlot
# Will plot box plot in ggplot of data simulated
#inputs: data frame, xVar, yVar, fillVar
#output the blox plot of data
#-----------------------------------------------
ANOVAplot <-function(data=df){
  names(data)<-c("Treatment","Value")
  plot<-ggplot(data=data,aes(x=Treatment,y=Value ,fill=Treatment)) + geom_boxplot()
return(plot)
}
ANOVAplot()
```
The above structured program is the proper way to set up an R Script for data analysis. As you can see, the libraries needed and global variables were names first. Then, I went through and wrote down all of my functions that I expect to be using in my program. From here I can then start my other coding and data analysis as needed
```{r}
#Calling my functions:
RandomData()
myANOVA()
ANOVAplot()
```
Now I am going to alter the functions so that they create a bar plot instead of a box plot:
```{r}
####################################
# Function: BRandomData
# These function will create random normal data to help simulate and test hypotheses.
#input: mean, n, standard deviation
#output: simulated random normal data set
#-----------------------------------------
BRandomData <-function(mean1=.6, mean2=.75,mean3=.95,mean4=1.1,n1=10,n2=10,n3=10, n4=10, sd1=.3, sd2=.3,sd3=.3,sd4=.3){
  myDF<-data.frame(Control=rnorm(mean=mean1, n=n1, sd=sd1), Light=rnorm(mean=mean2, n=n2, sd=sd2), Moderate=rnorm(mean=mean3, n=n3, sd=sd3), Extreme=rnorm(mean=mean4, n=n4, sd=sd4))
                  return(myDF)}
################################
#Function: BmyANOVA
#The function will perform an ANOVA analysis and spit generate p values
#input: data frame from 1st function
#output: p values analysis
#-----------------------------
BmyANOVA <- function(dataFrame=df){
 ANOVA <- aov(value~variable, data=dataFrame)
  myOutput <- summary(ANOVA)
  return(myOutput)
}

##############################
#Function: BPlot
# Will plot box plot in ggplot of data simulated
#inputs: data frame, xVar, yVar, fillVar
#output the blox plot of data
#-----------------------------------------------
BPlot <-function(data=df){
  plot<-ggplot(df, aes(x=variable, y=value, fill=variable)) +
  geom_bar(stat="identity")+theme_minimal() 
  return(plot)
}
```

PLotting both graphs side by side
```{r}
ANOVAplot()
BPlot()
```
I am now able to visualize the data in 2 different types of graph. I am still reading up on how to place error bars on the bar chart. I would also like to create a for loop of this too at some point to simulate the data a few times. I will try to get to that next week.